---
title: "Reproducible Research Project 1"
author: "Jingqin Luo"
date: "4/11/2021"
output: html_document
---


```{r setup, include=FALSE}
#knitr::opts_knit$set(root.dir = '//files.wustl.edu/jingqinluo/Reproducible Research')
knitr::opts_knit$set(root.dir = 'W:/Jingqin/1_ClinicalTools/Reproducible Research/')
knitr::opts_chunk$set(echo = TRUE,warning=FALSE, message=FALSE,cache=FALSE)
```
## Overview
This is my RR project 1 assignment of analyzing activitiy dataset while using *R* markdown to integrate R code, outputs and texts.  


## load in R packages
```{r load in packages}
library(RCurl)#for getting url
library(data.table)
library(dplyr)
library(tidyverse)
sessionInfo()##for reproducibility on R version and Packages
```

## Read in steps dataset
The steps data set was downloaded from the *Coursera* [course website](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip).

```{r read in step dat directly from github,echo=FALSE}
##try reading from github directly but failed.
#dat=read.table("https://github.com/rosyluo/RepData_PeerAssessment1/blob/master/activity.zip",sep=",",header=TRUE)#read directly from github

#url0 = getURL("https://github.com/rosyluo/RepData_PeerAssessment1/blob/master/activity.zip")
#dat = read.csv(text = url0)
```

```{r read in step dat}
##if download to local folder
temp=download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", destfile="activity.zip")#reduce manual downloading by directly downloading in R from url 
#dat=fread("activity.zip",header=T) ##directly read in the zip file return mosaic characters?
unzip("activity.zip")
dat=fread("activity.csv",header=T)
dim(dat)
str(dat)
head(dat)
tail(dat)
summary(dat)
```
## Calculate daily total steps, daily 5 minute interval mean and median and plot historgram
```{r process data for daily sum}
daySum = dat %>% group_by(date) %>% summarise(totalSteps=sum(steps,na.rm=T),meanSteps=mean(steps,na.rm=T),medianSteps=median(steps,na.rm=T))
dim(daySum)
print(daySum,type="html")
hist(daySum$totalSteps,bins=20,xlab="daily total steps", main="Histogram of daily total steps")
```

## Mean and median number of steps taken each day
The mean and median number of steps taken daily are as below.

```{r summarize daily sum: mean and median}
mean_median= daySum %>% summarise(mean=mean(totalSteps,na.rm=T),median=median(totalSteps,na.rm=T))
mean_median
```
## Histogram and time series plot of 5-min interval the average number of steps taken

```{r histogram of 5-minute interval steps before data imptuation}
hist(dat$steps,bins=100,xlab="5-minute steps", main="Histogram of daily total steps before imputation")
```
```{r time series plot}
summary(dat$steps)
mean0 = mean(dat$steps,na.rm=T)
plot(steps~date, data=dat,type="l")
abline(h=mean0, col= "red", lwd=2)
```

## The 5-minute interval that, on average, contains the maximum number of steps
```{r time interval with max steps}
dat[steps==max(steps,na.rm=T),]
```
The maximum steps taken in any 5-min interval is `max(steps,na.rm=T)`.

## describe and code a strategy for imputing missing data: using daily mean

```{r count NA}
count_NA = dat %>% summarise(total=n(), countNA=sum(is.na(steps)),fraction=countNA/total)
count_NA
```
There are `count_NA$countNA` total and  `count_NA$fraction` of 5-min intervals with missing data.

We will simply use the daily mean to impute the missing values in each 5-min interval.
```{r missing value imputation}
dayImputed = left_join(dat, daySum, by="date")
dayImputed$steps[which(is.na(dayImputed$steps))] = dayImputed$meanSteps[which(is.na(dayImputed$steps))]
head(dayImputed)
#print(dayImputed,type="html")
```
We replot the 5-min interval steps after missing data imputation.
## Histogram of the total number of steps taken each day after missing values are imputed

```{r histogram of 5-minute interval steps after data imptuation}
hist(dayImputed$steps,bins=100,xlab="5-minute steps", main="Histogram of daily total steps after imputation")
```
## Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r five minute summary by weekday vs weekend, echo=TRUE}
dayImputed$weekday = weekdays(dayImputed$date) 
dayImputed$weekday_flag = with(dayImputed, ifelse(weekday%in%c("Saturday","Sunday"),"weekend","weekday"))
mean_5minInterval = dayImputed %>% group_by(weekday_flag,interval)%>% summarise(mean=mean(steps,na.rm=T))
```

```{r panel plot on 5-minute mean inteval by weekday vs. weekend}
p = ggplot(dat=mean_5minInterval,aes(x=interval,y=mean))+geom_point()+geom_line()+
    facet_wrap(~weekday_flag)
p = p + theme_bw()+labs(x="time itnerval",y="averaged steps")+ theme(axis.text=element_text(size=15,face="bold"),axis.title=element_text(size=15,face="bold"),                                         legend.box=NULL,legend.title=element_text(size=15,face="bold"),legend.text=element_text(size=15,face="bold"),                      strip.background=element_rect(fill="white",colour="white"),strip.text=element_text(size=15,face="bold"))
plot(p)
```
As shown in the plots, during weekdays, the peak activities is during the interval 750~1000 only. On weekends, the first activity peak also appeared around 750~1000, but then the activities continued high for some periods from 1000~1750 .   
```{r extract to Rscript, echo=FALSE, include=FALSE}
#knit using R
#knit2html()
#to extract all the Rcode into a R script file, run in console: 
#setwd("W:/Jingqin/JosephIppolito/Lung cancer metabolite Sex difference/Rcode")
#knitr::purl("W:/Jingqin/JosephIppolito/Lung cancer metabolite Sex difference/Rcode/SexSpecificMetaboliteDifferenceLungCancer.Rmd")
```